Formalizing Stack Safety as a Security Property

### 2023-01-29 submission5/ideas-v2.tex
Work on key ideas

The paper is structured as follows. In \cref{sec:example}, we will walk through a function
call in a simple example machine and discuss how each of our properties applies to it,
informally. Then in \cref{sec:machine} we will introduce the machine model that we test, and give
its {\em security semantics}, described in brief below. We formalize the definitions of
stack safety that derive from the security semantics in \cref{sec:props}.
\Cref{sec:testing} describes the testing framework, and
\cref{sec:relwork,sec:future} discuss related and future work.
The remainder of this section introduces ideas that will be important throughout.

\paragraph*{Security Semantics and Property Structure}

@@

\paragraph*{The Properties and their Hierarchy}

\paragraph*{Events}

@@

\paragraph*{Threat Model}

@@

@@

\paragraph*{Permissiveness, Laziness, and Tagged Architectures}

This key motivation leads us to present integrity and confidentiality as predicates on (pairs of)
machine states that make claims about the future observable behavior of execution from those states.
That is not to say that our properties are specialized for lazy micro-policies. It also supports
normal ``eager'' micro-policies, and should support other enforcement mechanisms. In fact, \(\clri\)
and \(\clrc\) are even weaker than necessary to support Roessler and Dehon's micro-policy,
while \(\clec\) and \(\clei\) are stronger properties that they do not attempt to enforce.
%
\rb{Does this discussion (at least in part) belong in the general background
  section, some even in key ideas? It still feels a little out of order to me,
  and occasionally too detailed for the introduction?}

\rb{Begins to anticipate subsequent discussion. This is even something of an
  understatement: some papers (esp. newer ones?), make concrete claims, though
  they remain informal. Arguably, none have a truly formal characterization in
  mind (with the qualified exception of \cite{SkorstengaardSTKJFP}): this is one
  of the novelties of this work!
\rb{(+SNA) Indeed, this point needs to be made carefully. We claim to go further
  than previous work, without showing a fully general framework (which would be
  highly parametric, etc.). Tie to the idea of micro-policies as a good place to
  start because of their flexibility, elegance, enabling a wide of interesting
  experiments, etc.
\rb{Basically, none reach the point of being an independent (and executable) specification}
}
}
%
Previous attempts
to produce a formal characterization do so using a {\it fully abstract overlay semantics},
a style of correct-by-construction machine \cite{SkorstengaardSTKJFP}.
But this approach is not suitable for the {\em lazy} enforcement mechanism
of Roessler and DeHon, in which stores are not checked, and illegal stores are
only caught when their location is later read. We aim to provide a stack-safety
property suitable for testing lazy enforcement.
%
\rb{Need to discuss overlay semantics, but also note that it is very intensional
  and very specific, as the authors acknowledge. I'd introduce this a bit more
  gradually. Also, if phrased this way, it sounds like its main drawback is a
  lack of generality, to which we should contrast a more general approach, not
  just one that is specialized to a specific policy, i.e. lazy micro-policies
  (I'd try to present these more gradually as well).}

\begin{table}
\begin{threeparttable}
  \begin{tabular}{l | c c c c | c c c | c}
    \toprule
    \multicolumn{1}{l |}{Mechanism}
    & \multicolumn{4}{l |}{Features}
    & \multicolumn{3}{l |}{Properties}
    & %% Formal criteria
    \\
    \midrule
    %% Technique
    \multicolumn{1}{c}{}
    & \rot{Strict enforcement}
    & \rot{Deferred checking}
    & \rot{Low-level code support}
    & \multicolumn{1}{c}{\rot{Partial program support}}
    & \rot{Well-bracketed control flow}
    & \rot{Data integrity}
    & \multicolumn{1}{c}{\rot{Data confidentiality}}
    & \rot{Formal criteria}
    \\
    \midrule
    Stack canaries  & \tabno   & \tabjein & \tabyes  & \tabyes  & \tabyes  & \tabno   & \tabno   &          \\
    Shadow stacks   & \tabno   & \tabjein &          & \tabyes  & \tabyes  & \tabno   & \tabno   &          \\
    Bounds checking & \tabyes  & \tabno   & \tabno   & \tabno   & \tabyes  &          &          &          \\
    Split stacks    & \tabyes  & \tabno   & \tabno   & \tabyes  & \tabyes  & \tabno   & \tabno   &          \\
    \midrule
    Capabilities    &          & \tabno   &          & \tabyes  &          &          &          & \tabjein \\
    Micro-policies  & \tabyes  & \tabyes  & \tabyes  & \tabyes  & \tabyes  & \tabyes  & \tabyes  & \tabyes  \\
    \bottomrule
  \end{tabular}
  \caption{Comparison of Stack Protection Mechanisms}
  \begin{tablenotes}
  \item \tabyes indicates a protection mechanism possesses the feature
  \item \tabno indicates the mechanism lacks the feature
  \item \tabjein indicates partial or incomplete support
  \item \rb{Complete and check columns, describe. Capitalize?}
  \item \rb{Order? Split int/conf? Deferred/lazy at the end of features}
  \item \rb{2+ lines for capabilities, before and after using our draft?}
  \end{tablenotes}
  \label{tab:relwork}
\end{threeparttable}
\end{table}

@@

We have made explicit the key point that there is \emph{not} a single stack
abstraction. This paves the way for a \emph{general} idea of stack protection
that is widely applicable with high-level concepts that remain unchanged (and I
think it's particularly important to at least introduce these more immutable
parts early on). The fact that this general ``framework'' requires low-level or
specific specializations is a predictable consequence of the protean nature of
the call stack.
Let us return to a selection of representative policies in our menagerie and
consider how they fit into the framework. At the highest level, we can classify
all those techniques into two large groups, software-based and hardware-based.
Traditionally, most practical protections are implemented in software, so we
start by considering this first group.
We can make two general observations about these software based techniques.
Firstly, (as we will see: discussion? technical details?) these techniques are
\emph{less modular} than their hardware counterparts. That is, the
countermeasures are tightly coupled with the program code they protect, and
consequently more effort is required to disentangle the program from the
protections, and to understand the impact of the latter on the former. This is
true of any systematic attempt to study the security of stack protection
policies, including their adaptation to our framework. (Owing to this added
complexity, software policies will not be the cleanest or most elegant exemplars
of use of the framework, its salient features and practical utility; this partly
motivates the focus on hardware policies.)
(The above remark holds \emph{at least in a ``single-level model''} that
considers assembly language/machine code programs. In general, (low-level)
protections are introduced by a compiler or similar tool. One could consider a
multi-level extension of the framework pairing models of high-level programs and
the security transformations performed by a (stack-secure) compiler.)
Secondly, the guarantees generally offered by software-based policies are
\emph{less absolute}, being defined more or less informally and usually
providing few definite guarantees of absence of \emph{general} classes of
errors. (The following discussion elaborates on this. It is conceivable that an
extended and more sophisticated framework could add \emph{probabilistic
reasoning} capabilities to capture and reason about the effects of less strict
protection mechanisms. This would be another orthogonal development to what is
presented here. In this paper, we are interested in assessing the complete
\emph{absence} of broad classes of vulnerabilities --- modulo any modeling gaps
or defects. In passing, this would extend the security properties to e.g. larger
system verification efforts.)
One by one, consider the following specimens of software protection mechanisms
(quick explanations inlined?):
\begin{itemize}
\item \emph{Stack canaries}. The purpose of canaries is to \emph{increase} the
  difficulty of some attacks on the stack, usually (always?) buffer overflows.
  They can never provide any absolute security guarantees and therefore do not
  satisfy any of our properties (some of the properties, e.g. confidentiality,
  are completely out of scope.) Although tedious, we could we could model this
  mechanism in our framework with the appropriate adaptations --- and
  unsurprisingly, testing would find property violations, which we already know
  are there: the results of this effort would be unsurprising and not as useful
  or instructive as other experiments.
  \rb{(+SNA) They are kind of lazy, aren't they?}
\item \emph{Bounds checking}. This is fundamentally a high-level technique, even
  more obviously so than stack canaries (which could still detect violations
  caused by external code). Bounds checking only applies to code that has been
  instrumented by the ``secure compiler'' (i.e., to whole programs); it does not
  impose any constraints on adversarial code. The attacker model is therefore
  rather weak. (Certain variants, like SoftBound, are good targets for lifting
  our framework to the source level by e.g. using a C-language semantics and
  using Tagged C for security modeling; the lifted versions of our properties
  would likely hold under the expected assumptions. HardBound is probably an
  interesting transitional case because it uses capability-like hardware
  primitives for enforcement; it could be close to ongoing and future work to
  verify stack safety on capability machines, e.g. based on Cerise -- discuss
  together?)
\item \emph{Shadow stacks}. Like stack canaries, shadow stacks are
  ``probabilistic techniques'' in that they ``restrict the flexibility available
  in creating gadget chains'' (Shanbhogue et al.). This technique is solely
  concerned with control-flow integrity, and so closely related to our
  well-bracketed control flow. Nevertheless, and again, it is only an
  approximate technique, and our testing framework would be able to find the
  expected attacks against our security property.
\item \emph{Split stacks}. The security claims of split stacks are stronger than
  those of most software countermeasures, i.e., they are meant to guarantee
  their titular code-pointer integrity. They explicitly offer no protections on
  stack data (i.e., integrity and confidentiality), and so restrict themselves
  to well-bracketed control flow. (This is probably closer to our framework than
  most other software techniques, and it seems like it would be easier to model
  as well. However, this is again work based on high-level languages and would
  benefit from a higher level of abstraction.)
\item \ldots
\end{itemize}
Opposed to the above group of software techniques we have hardware-based
protection policies (more in background/technical preliminaries?). Although
programs must effectively leverage the protection mechanisms offered by the
hardware, the distinction between those protections and the program logic itself
is much cleaner: it is fairly simple to separate the use of the security
features in the hardware (and their effects in program code) from regular
program code. They also tend to aspire to offer complete protection against
entire classes of attacks. Both aspects make this family of techniques a better
fit for illustrating and showcasing our techniques.

@@

\rb{(+SNA) Any interesting tables comparing specimens in the menagerie?}
\rb{(+SNA) A general multi-level extension of the framework would need to model
  multiple source languages. Among other things, the stack is the meeting place
  of all sorts of different language abstractions! One of the advantages of the
  single-level model is that it lets us be agnostic about the language. (On the
  other hand, it is likely that a single source and target mediated by the
  compiler suffice to model many existing techniques, coming e.g. from systems
  programming and C-like languages.}

\paragraph*{Limitations}

Our properties are termination insensitive: in order to support enforcement mechanisms
that failstop rather than execute dangerous code, the properties must treat failstops
as invisible to the attacker. We extend this insensitivity to all errors for simplicity.
This is consistent with our treatment of properties that specify the state after a return,
in that a callee that diverges for any reason will never reach the return, and will fulfill
the property vacuously.
%
\rb{Limitations in the introduction or later? Orthogonality of the heap, morally
  similar interesting setups on the stack? Termination sensitivity and
  capability machine semantics?}

@@

>>>>

The paper is structured as follows. In \cref{sec:example}, we will walk through a function
call in a simple example machine and discuss how each of our properties applies to it,
informally. Then in \cref{sec:machine} we will introduce the machine model that we test, and give
its {\em security semantics}, described in brief below. We formalize the definitions of
stack safety that derive from the security semantics in \cref{sec:props}.
\Cref{sec:testing} describes the testing framework, and
\cref{sec:relwork,sec:future} discuss related and future work.
The remainder of this section introduces ideas that will be important throughout.
%
%% \rb{Or shorter: references in contributions + final sentence at end of intro}

%% \rb{Turn these back into paragraphs later?}
\subsection{Permissiveness, Laziness, and Tagged Architectures}

This key motivation leads us to present integrity and confidentiality as predicates on (pairs of)
machine states that make claims about the future observable behavior of execution from those states.
That is not to say that our properties are specialized for lazy micro-policies. It also supports
normal ``eager'' micro-policies, and should support other enforcement mechanisms. In fact, \(\clri\)
and \(\clrc\) are even weaker than necessary to support Roessler and Dehon's micro-policy,
while \(\clec\) and \(\clei\) are stronger properties that they do not attempt to enforce.
%
\rb{Emphasize the point about eager properties, here, when discussing other
  enforcement types?}
%
\rb{(Mentioned properties are still ``forward references'' of a sort as they
  haven't been formally introduced yet)}
%% \rb{Does this discussion (at least in part) belong in the general background
%%   section, some even in key ideas? It still feels a little out of order to me,
%%   and occasionally too detailed for the introduction?}

\rb{Moved last bit to the framework discussion below, does this work?}

\rb{Test: the comparison of enforcement mechanisms \emph{and} the hierarchy of
  properties \emph{before} the security semantics. Does this work? Easy to move
  around/consolidate. Some things still feel a bit out of order. The comparison
  motivates the hierarchy, and also parts of the security semantics -- too much
  at once? Also: bad subsection/paragraph title ahead.}

\subsection{A Formal and Flexible Security Framework}

While a concrete aim is to provide a stack-safety property suitable for testing
lazy enforcement for a micro-policy architecture, the formal framework we
develop could be adapted to other protection mechanisms.
%
The need for these policy-specific adaptations is a natural and predictable
consequence of the protean nature of the call stack (discussed in \cref{XXX}),
and does not detract from the generality of our properties and their
underlying \rb{high-level?} design principles.
%
Nevertheless, some enforcement methods lends themselves to modeling more easily
than others. Micro-policies in particular stand out for their elegance, which
results in particularly clean instances of our properties, and their
flexibility, which enables them to implement very general forms of stack
protection \rb{(and so we get to exercise our properties fully)}.

Let us revisit our menagerie of stack protections from \cref{XXX} to motivate
our choices. \cref{tab:relwork} summarizes and compares those techniques along
several representative axes.

\begin{table}
\begin{threeparttable}
  \begin{tabular}{l | c c c c | c c c | c}
    \toprule
    \multicolumn{1}{l |}{Mechanism}
    & \multicolumn{4}{l |}{Features}
    & \multicolumn{3}{l |}{Properties}
    & %% Formal criteria
    \\
    \midrule
    %% Technique
    \multicolumn{1}{c}{}
    & \rot{Strict enforcement}
    & \rot{Deferred checking}
    & \rot{Low-level code support}
    & \multicolumn{1}{c}{\rot{Partial program support}}
    & \rot{Well-bracketed control flow}
    & \rot{Data integrity}
    & \multicolumn{1}{c}{\rot{Data confidentiality}}
    & \rot{Formal criteria}
    \\
    \midrule
    Stack canaries  & \tabno   & \tabjein & \tabyes  & \tabyes  & \tabjein & \tabno   & \tabno   &          \\
    Shadow stacks   & \tabno   & \tabjein &          & \tabyes  & \tabjein & \tabno   & \tabno   &          \\
    Bounds checking & \tabyes  & \tabno   & \tabno   & \tabno   & \tabyes  &          &          &          \\
    Split stacks    & \tabyes  & \tabno   & \tabno   & \tabyes  & \tabyes  & \tabno   & \tabno   &          \\
    \midrule
    Capabilities    &          & \tabno   &          & \tabyes  &          &          &          & \tabjein \\
    Micro-policies  & \tabyes  & \tabyes  & \tabyes  & \tabyes  & \tabyes  & \tabyes  & \tabyes  & \tabyes  \\
    \bottomrule
  \end{tabular}
  \caption{Comparison of Stack Protection Mechanisms}
  \begin{tablenotes}
  \item \tabyes indicates a protection mechanism possesses the feature
  \item \tabno indicates the mechanism lacks the feature
  \item \tabjein indicates partial or incomplete support
  \item \rb{Complete and check columns, describe. Capitalize?}
  \item \rb{Order? Split int/conf? Deferred/lazy at the end of features}
  \item \rb{2+ lines for capabilities, before and after using our draft?}
  \end{tablenotes}
  \label{tab:relwork}
\end{threeparttable}
\end{table}

\subsubsection{Strict Property Enforcement}
%
We are primarily interested in the \emph{strict} enforcement of a set of
security properties of choice, i.e., no violations are allowed. Some \rb{most?}
of the most commonly deployed countermeasures, like stack canaries and shadow
stacks, offer only \emph{approximate} enforcement. As such, they will not
satisfy any of our properties per se. We could use our framework to model them
with some effort, but the interest of this exercise would be limited: testing
would easily find the expected, unavoidable property violations. A more complex,
\emph{probabilistic} version of the framework could offer adequate reasoning
capabilities for these defenses.

\subsubsection{Deferred and Lazy Properties}
%
We seek to support both \emph{eager} property enforcement (where potential
violations are detected immediately) as well as more permissive \emph{lazy}
properties (as motivated in \cref{XXX}).
%
Micro-policies are well suited to both types of properties. There is an
interesting connection between lazy enforcement and the deferred checks of stack
canaries and shadow stacks. Other techniques, like split stacks and bounds
checking---and the closely related hardware capabilities---are fundamentally
eager techniques.

\subsubsection{Low-Level Component Support}
%
Our framework considers a \emph{single-, low-level view} of all program code and
data; a key advantage of this approach is that it allows us to remain
\emph{language-agnostic}.
%
Some software-based protection mechanisms are designed specifically with
a \emph{high-level language} in mind, e.g. bounds checking and split stacks.
They tend to offer stronger security guarantees than other software techniques.
In exchange, their threat model becomes weaker, as they make additional
assumptions about the admissible behavior of program components
%
(for example, bounds checking only applies to code that has been instrumented by
a \emph{secure compiler} \ref{}).
%
In general, a multi-level extension of the framework offering adequate reasoning
capabilities for this class of policies would need to model multiple source
language and a notion of code provenance---a reflection of the call stack as the
melting pot of language abstractions.
%
The common case of one source and one target language mediated by a compiler may
be solved by a simpler lifting of our framework from the machine to the source
level \rb{e.g., SoftBound could use a C language semantics and Tagged C for
security modeling (lifted properties would likely hold under the expected
assumptions)}.

\subsubsection{Partial Program Protection}
%
A consequence of the use of high-level software techniques is the difficulty of
keeping their shields up against a \emph{low-level attacker} that is not bound
to play by their rules.
%
Insofar as their protections continue to work against untrusted code, like split
stacks, they do so by relying on additional support, e.g. from the operating
system.
%
\rb{(The same remarks about the language-agnostic framework apply here. Split
  stacks are probably relatively easy to model without big changes, but would
  still benefit from a higher level of abstraction.)}
%
\rb{(partial programs and split stacks? jein?)}

\subsubsection{Enforced Properties}
%
\rb{(FIXME: Present with the hierarchy of properties? Basically all policies try
  to protect {\wbcf}. Without it things are quite strange. Some of the more
  restrictive software methods and the hardware methods can cover additional
  properties.)}

\subsubsection{Software vs. Hardware}
%
One first general observation about software-based techniques is that they
are \emph{less modular} than their hardware counterparts, that is, the
countermeasures are more tightly coupled with the program code they protect.
%
By contrast, although programs must effectively leverage the protection
mechanisms offered by the hardware, the distinction between those protections
and the program logic itself is much cleaner.
%
Our framework relies on a \emph{separation} of security-relevant operations and
regular program code \rb{stress that any scalable, composable solution must do
this!}.
%
While the various extensions suggested in this section could help abstract the
complexities of software safeguards, hardware techniques remain the most elegant
and general exemplars for the methods we seek to demonstrate.
%
\rb{(This is also related to our single-level model)}

\subsubsection{Formal Models}
%
A second general observation about software-based techniques is that they tend
to be \emph{less absolute}: they enforce weaker properties or under stronger
assumptions, and their threat models are usually fairly informal.
%
%
The closest previous attempt
to produce a formal characterization of stack safety does so using a {\em fully abstract overlay semantics},
a style of correct-by-construction machine \cite{SkorstengaardSTKJFP}.
But this approach
builds the security properties into the semantics of the machine without externalizing them or their meaning.
It is also not suitable for the {\em lazy} enforcement mechanism
of Roessler and DeHon, in which stores are not checked, and illegal stores are
only caught when their location is later read.
%
\rb{Need to discuss overlay semantics, but also note that it is very intensional
  and very specific, as the authors acknowledge. I'd introduce this a bit more
  gradually. Also, if phrased this way, it sounds like its main drawback is a
  lack of generality, to which we should contrast a more general approach, not
  just one that is specialized to a specific policy, i.e. lazy micro-policies
  (I'd try to present these more gradually as well).}
%
\rb{Also, testing certainly (and motivate), but not only?}

Overall, the generality and flexibility of micro-policies, arising from an
elegant design that clearly separates the program from its security, are the
most appropriate choice for an exploration of this new framework.
%
\rb{FIXME: Finish motivation. We focus on micro-policies, but\ldots Something
  about existing policies, exercising the framework most generally,
  representative experiments?}

\rb{Begins to anticipate subsequent discussion. This is even something of an
  understatement: some papers (esp. newer ones?), make concrete claims, though
  they remain informal. Arguably, none have a truly formal characterization in
  mind (with the qualified exception of \cite{SkorstengaardSTKJFP}): this is one
  of the novelties of this work!
\rb{(+SNA) Indeed, this point needs to be made carefully. We claim to go further
  than previous work, without showing a fully general framework (which would be
  highly parametric, etc.). Tie to the idea of micro-policies as a good place to
  start because of their flexibility, elegance, enabling a wide of interesting
  experiments, etc.
\rb{Basically, none reach the point of being an independent (and executable) specification}
}
}

\subsection{The Properties and Their Hierarchy}

\subsection{Security Semantics and Property Structure}

@@

\subsection{Events}

@@

\subsection{Threat Model}

@@

\subsection{Limitations}

Our properties are termination insensitive: in order to support enforcement mechanisms
that failstop rather than execute dangerous code, the properties must treat failstops
as invisible to the attacker. We extend this insensitivity to all errors for simplicity.
This is consistent with our treatment of properties that specify the state after a return,
in that a callee that diverges for any reason will never reach the return, and will fulfill
the property vacuously.
%
\rb{Merge with threat model? Orthogonality of the heap, morally
  similar interesting setups on the stack? Termination sensitivity and
  capability machine semantics?}

@@

@@

%% We have made explicit the key point that there is \emph{not} a single stack
%% abstraction. This paves the way for a \emph{general} idea of stack protection
%% that is widely applicable with high-level concepts that remain unchanged (and I
%% think it's particularly important to at least introduce these more immutable
%% parts early on). The fact that this general ``framework'' requires low-level or
%% specific specializations is a predictable consequence of the protean nature of
%% the call stack.
%% %
%% \\
%% %
%% Let us return to a selection of representative policies in our menagerie and
%% consider how they fit into the framework. At the highest level, we can classify
%% all those techniques into two large groups, software-based and hardware-based.
%% Traditionally, most practical protections are implemented in software, so we
%% start by considering this first group.
%% %
%% \\
%% %
%% We can make two general observations about these software based techniques.
%% Firstly, (as we will see: discussion? technical details?) these techniques are
%% \emph{less modular} than their hardware counterparts. That is, the
%% countermeasures are tightly coupled with the program code they protect, and
%% consequently more effort is required to disentangle the program from the
%% protections, and to understand the impact of the latter on the former. This is
%% true of any systematic attempt to study the security of stack protection
%% policies, including their adaptation to our framework. (Owing to this added
%% complexity, software policies will not be the cleanest or most elegant exemplars
%% of use of the framework, its salient features and practical utility; this partly
%% motivates the focus on hardware policies.)
%% %
%% \\
%% %
%% (The above remark holds \emph{at least in a ``single-level model''} that
%% considers assembly language/machine code programs. In general, (low-level)
%% protections are introduced by a compiler or similar tool. One could consider a
%% multi-level extension of the framework pairing models of high-level programs and
%% the security transformations performed by a (stack-secure) compiler.)
%% %
%% \\
%% %
%% Secondly, the guarantees generally offered by software-based policies are
%% \emph{less absolute}, being defined more or less informally and usually
%% providing few definite guarantees of absence of \emph{general} classes of
%% errors. (The following discussion elaborates on this. It is conceivable that an
%% extended and more sophisticated framework could add \emph{probabilistic
%% reasoning} capabilities to capture and reason about the effects of less strict
%% protection mechanisms. This would be another orthogonal development to what is
%% presented here. In this paper, we are interested in assessing the complete
%% \emph{absence} of broad classes of vulnerabilities --- modulo any modeling gaps
%% or defects. In passing, this would extend the security properties to e.g. larger
%% system verification efforts.)
%% %
%% \\
%% %
%% One by one, consider the following specimens of software protection mechanisms
%% (quick explanations inlined?):
%% %
%% \\
%% %
%% \begin{itemize}
%% %
%% \item \emph{Stack canaries}. The purpose of canaries is to \emph{increase} the
%%   difficulty of some attacks on the stack, usually (always?) buffer overflows.
%%   They can never provide any absolute security guarantees and therefore do not
%%   satisfy any of our properties (some of the properties, e.g. confidentiality,
%%   are completely out of scope.) Although tedious, we could we could model this
%%   mechanism in our framework with the appropriate adaptations --- and
%%   unsurprisingly, testing would find property violations, which we already know
%%   are there: the results of this effort would be unsurprising and not as useful
%%   or instructive as other experiments.
%%   \rb{(+SNA) They are kind of lazy, aren't they?}
%% %
%% \item \emph{Bounds checking}. This is fundamentally a high-level technique, even
%%   more obviously so than stack canaries (which could still detect violations
%%   caused by external code). Bounds checking only applies to code that has been
%%   instrumented by the ``secure compiler'' (i.e., to whole programs); it does not
%%   impose any constraints on adversarial code. The attacker model is therefore
%%   rather weak. (Certain variants, like SoftBound, are good targets for lifting
%%   our framework to the source level by e.g. using a C-language semantics and
%%   using Tagged C for security modeling; the lifted versions of our properties
%%   would likely hold under the expected assumptions. HardBound is probably an
%%   interesting transitional case because it uses capability-like hardware
%%   primitives for enforcement; it could be close to ongoing and future work to
%%   verify stack safety on capability machines, e.g. based on Cerise -- discuss
%%   together?)
%% %
%% \item \emph{Shadow stacks}. Like stack canaries, shadow stacks are
%%   ``probabilistic techniques'' in that they ``restrict the flexibility available
%%   in creating gadget chains'' (Shanbhogue et al.). This technique is solely
%%   concerned with control-flow integrity, and so closely related to our
%%   well-bracketed control flow. Nevertheless, and again, it is only an
%%   approximate technique, and our testing framework would be able to find the
%%   expected attacks against our security property.
%% %
%% \item \emph{Split stacks}. The security claims of split stacks are stronger than
%%   those of most software countermeasures, i.e., they are meant to guarantee
%%   their titular code-pointer integrity. They explicitly offer no protections on
%%   stack data (i.e., integrity and confidentiality), and so restrict themselves
%%   to well-bracketed control flow. (This is probably closer to our framework than
%%   most other software techniques, and it seems like it would be easier to model
%%   as well. However, this is again work based on high-level languages and would
%%   benefit from a higher level of abstraction.)
%% %
%% \item \ldots
%% %
%% \end{itemize}
%% %
%% Opposed to the above group of software techniques we have hardware-based
%% protection policies (more in background/technical preliminaries?). Although
%% programs must effectively leverage the protection mechanisms offered by the
%% hardware, the distinction between those protections and the program logic itself
%% is much cleaner: it is fairly simple to separate the use of the security
%% features in the hardware (and their effects in program code) from regular
%% program code. They also tend to aspire to offer complete protection against
%% entire classes of attacks. Both aspects make this family of techniques a better
%% fit for illustrating and showcasing our techniques.
%% %
%% \\
%% %

@@

%% \rb{(+SNA) Any interesting tables comparing specimens in the menagerie?}
%% \rb{(+SNA) A general multi-level extension of the framework would need to model
%%   multiple source languages. Among other things, the stack is the meeting place
%%   of all sorts of different language abstractions! One of the advantages of the
%%   single-level model is that it lets us be agnostic about the language. (On the
%%   other hand, it is likely that a single source and target mediated by the
%%   compiler suffice to model many existing techniques, coming e.g. from systems
%%   programming and C-like languages.}

@@

\rb{Deferred/lazy connection of canaries and shadow stacks -- arguably these helps them support things other software policies need more help with?}

\rb{Capabilities and HardBound?}
